---
title: "Employee Attrition EDA for DDSAnalytics"
author: "Dustin Bracy"
date: "12/05/2019"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(kableExtra)
```

## R Markdown

Objectives:  
- Identify the top three factors that contribute to employee attrition (backed up by evidence provided by analysis).
- Identify job and/or role specific trends.
- Identify useful factors related to talent management (workforce planning, employee training, identifying high potential employees, reducing turnover)
- Present findings via YouTube in 7 minutes or less

Target Audience:
- CEO and CFO of Frito Lay
- CEO is statistician, CFO has had only one class in statistics

Clarification:
- Job level 1-5 low-high
- Overtime = non-exempt
- Distance is unknown (mi/km), suggest using high/low
- use percent attrition vs count
- performance ratings maybe indicate there is a fear of giving a low score
- training times last year = # of traning sessions attended
- what is hourly/daily/monthly rate?  Production rates?  

```{r}
data <- read.csv('./data/CaseStudy2-data.csv', header=T)
summary(data)

# Check for missing values:
MissingValues <- sapply(data, function(x)sum(is.na(x)))
MissingValues %>% kable("html") %>% kable_styling()

data %>% group_by(Attrition) %>% summarise(Employees = n()) %>% ggplot(mapping=aes(AttritionCount, Employees)) + geom_col()


```
